\sys: Light-touch Data-Parallel Shell Processing

### 2020-05-22 pldi/draft.tex
Pass over related work

\heading{Discussion}
\sys seems to achieve near-linear speedup on low-parallelism configurations.
This is due to a high degree of task-based parallelism from distinct \emph{map} and \emph{fold} processes, which results in a longer tree of task-parallel stages.
This is seen in all that include components in \pur.
For example, Sort in $8\times$ spawns 37 nodes:
  8 \ttt{tr} nodes, 8 \ttt{sort} nodes, 7 \ttt{sort -m} nodes, and 14 \ttt{relay} nodes.
In effect, \sys hits the maximum number of parallel processes (and thus achieves optimal performance, \ie one without unnecessary process spawns and context switches) 
significantly earlier than the purported parallelism---$16-32\times$ for a 64-core system.

Thus, asking for parallelism beyond that point results in worse speedup, due to other overheads that start dominating as no further speedup is available.
Sort-sort with $32\times$ has 314 nodes (and 54m sys time) and with $64\times$ has 634 nodes (and 90m! sys time).
 On the other hand opt-bigrams 64 has 255 nodes and the sys time stays
  the same so I think that it might have to do with the fact that
  constant costs rise a lot (since opt-bigrams 32 takes just 1
  minute).

@@

%% \TODO{An important point that needs to be made clear is that these are
%%   taken from the wild unmodified. And that Pash never leads to
%%   slowdown (except for the trivial head cases, where the slowdown is
%%   neligible).}
\heading{Results}
\sys's parallelism was set to $16\times$, as many of these pipelines are relatively long (\S\ref{ours}, Discussion).
Their input was generated by multiplying the original Unix50 input several times to reach 10GB.
Fig~\ref{fig:unix50-individual} shows the speedup (left) over the sequential runtime (right) for all 34 pipelines.
The average speedup is $5.49$ ($6.07\times$), and the weighted average (with the absolute times as weights) is $5.75\times$.
 % the \todo{geometric mean is $1.08\times$},

@@

\heading{Discussion}
Most pipelines see significant acceleration---except 13, 24, 25, 26, 29, and 30 that see no speedup and 2, 19, and 31 that see a slowdown.
Ones in the first group contain stages that cannot be parallelized in general---\eg \ttt{awk}, \ttt{sed} \ttt{-d}, and \ttt{tr} \ttt{-d}.
However, by applying trivial changes
that change some of these commands with their equivalents in \sta---\eg, replacing a general \ttt{awk} with more specialized \ttt{cut} that achieves the same effect---\sys 
can identify and exploit parallelism:
  for, say, \todo{13} this leads  to \todo{XX}$\times$ speedup.
This is a common pattern we witnessed elsewhere too~\sx{macro1}.

Ones in the second group contain a \ttt{head}, thus always processing only one line and executing within $10ms$ regardless of the full input.
\sys's slowdown is due to constant setup costs---\eg pipes, processes---but remains within $1s$.

For pipelines  \sys performs the same as the sequential implementation.

@@

Finally, for the rest of the pipelines \sys speedup is \todo{capped}
because of a combination of the following reasons: (i) they contain
pure commands that are parallelizable but don't scale linearly, such
as \ttt{sort} (e.g. pipelines 0, 1, 3, 15, 16, 18, 20, 27, 28, 33),
(ii) they are deep pipelines that already exploit task-based
parallelism (e.g. pipelines 7, 8, 9, 10, 11, 12, 14, 27, 28), and
(iii) they are not cpu-intensive, leading to \todo{larger IO and
  constant costs} (e.g. pipelines 4, 5, 6, 7, 8, 12, 14, 22, 23).

@@

\heading{Take-aways} \sys leads to significant performance
improvements in completely unmodified \unix pipelines found out in the
wild. Small tweaks can yield further improvements, showing that
\sys-awareness and scripting expertise can improve
results. Furthermore, it doesn't cause slowdown in any non-trivial
computation, even in cases where parallelism is not possible.

@@

\heading{Program}
The full version of this program is inspired by the introductory chapter of Hadoop's Definitive Guide~\cite[Chapter 2]{hadoop:15}, where it is used as an example of a realistic mid-scale analytics pipeline.

The original task, as presented in the book, is comprised of three sub-tasks:
  download weather data from NCDC (shell), convert them to a format that is Hadoop-friendly (shell), and process them to identify the maximum temperature (Hadoop).
Only the last of these tasks is the real focus of the book, demonstrating the benefits of using distributed framework such as Hadoop.
For \sys, we consider the full pipeline that implements all three sub-tasks---from fetching data to outputting the maximum temperature.

The full program used for our evaluation is slightly different than the one presented in Fig.~\ref{fig:example}, consisting of 14 stages and totalling 240 characters.
We full program expresses the \ttt{for} loop as its first pipeline stage and uses \ttt{xargs} to 
The program presented in Fig.~\ref{fig:example} is simplified only for clarity of exposition.

\heading{Results}
% The first interesting result 
\TODO{Some results}
The Hadoop program, expressed in Java and only capturing the \emph{four} last stages, amounts to 137 lines of Java code---not accounting for commands moving data in and out of HDFS.

As this is benchmark is taken directly from the Hadoop book, it is worth reporting on its execution time.
While the setup is not typical of Hadoop, the comparison between the two systems is not unreasonable:  % we took several steps to make the comparison fair:
  (i) both Hadoop and Spark favor message-passing rather than shared memory,
	(ii) Hadoop has no fault-tolerance for master and worker nodes (\ie both systems do not replicate data),
	(iii) the task at hand is a good fit for  Hadoop,
	(iv) the Hadoop program is written as a single program tailored for this task---with ample opportunity for optimization from the compiler and runtime system---rather than a series of loosely coupled programs each written to handle many other cases.
\todo{The original pipeline executes in XXs---long enough to amortize Hadoop's 0.5--0.9s startup costs.}
With two Hadoop 3.2.1 (Openjdk 11.0.5) data-nodes, the Hadoop program takes 6m22s for data processing and 1m09s for moving the dataset to HDFS.
For the same amount of parallelism, \sys takes 4m30.554s---this includes downloading and uncompressing the data files which, as described earlier, is not expressible in the Hadoop equivalent.

@@

\heading{Discussion}
Similar to Unix50~\sx{unix50}, we found that large, complex pipelines enable significant freedom in terms of expressiveness.
Several stages of the original program were expressed as a single \ttt{awk} script;
  unfortunately, \ttt{awk} is too general to have a single meaningful \todo{parallelizability} signature.
A simpler (but longer) pipeline that leverages on \unix built-ins might be outperformed by a specialized programs on a single node, but turns out to be trivially scalable to multiple nodes.
% For example, the \ttt{sort} command at the end is an expensive version of extracting the maximum element;
%   this could have been expressed with a simple \ttt{awk} script that only keeps track of the max.

\heading{Take-aways}
Multi-line Java Program parallel program decomposed in two phases---a highly parallel map phase and a sequential reduce phase---corresponds to a much smaller shell script with

@@

\heading{Program} At a high level, the pipeline is comprised of two
distinct components that coarsely correspond to a front- and a
back-end communicating via \unix pipes.  The front-end downloads pages
and extracts their URLs from their source, feeding back \todo{to} the
start of the pipeline; the back-end extracts the text from the html
pages and then applies natural-language processing---\eg trigrams,
character conversion, term frequencies---to index the incoming pages.
The pipeline contains a total of 52 stages (18 for the font-end and 34
for the back-end) written in multiple languages---for example,
its \ttt{url-extract}ion utility is written in JavaScript whereas its
\ttt{word-stem}ming utility is in Python.  \sys can still operate on
them as their parallelizability properties---\sta for
\ttt{url-extract} and \ttt{word-stem}---can be described by its DSL.

\heading{Results} To simplify execution (and avoid hitting Wikipedia),
we have saved a recent version of Wikipedia to a server within the
network and to redirect all \ttt{curl} requests there.  The complexity
of this pipeline makes end-to-end measurements difficult, so we
measure the improvements on the two components components individually
and then report the differences. With parallelism set to \todo{2/16}
the execution time achieved by \sys is \todo{X} for the front-end and
\todo{Y} for the back-end, while the original execution time is
\todo{X} for the front-end and \todo{Y'} for the back-end. \kk{If the
  front-end takes a lot of time (compared to the back-end), we might
  better not report it with the premise that it is a feedback cycle
  that cannot be parallelized by \sys.}

\heading{Discussion}

Since the front-end contains a feedback cycle, \sys \todo{doesn't
  parallelize it}, and in both cases it takes \todo{X seconds} to
complete. However, \sys parallelizes the back-end and achieves a
speedup of \todo{X} with parallelism set to \todo{2/16}. Several
stages of the pipeline are \sta (such as the \ttt{word-stem} script
and the text extraction) and that allows \sys to achieve benefits by
exposing data parallelism. Note that the back-end part of the script
contains 34 stages executing in parallel, so the sequential version
\todo{already benefits from task-based parallelism} \kk{not totally
  true}.

@@

\subsection{Use Case: Container Orchestration}
\label{macro3}

As a third realistic workload, we use a script for orchestrating containers to test the latest built across various Linux distributions.
Compatibility testing is a common task for large software projects, commonly automated through shell scripting.

\heading{Program}
The 30-line orchestration script is unusual among our set of benchmarks in that it is not structured as a linear pipeline.
Rather, it makes extensive use of POSIX constructs common in the shell, including output redirection, conditionals, and here-documents.
It also combines commands from other classes outside \sta and \pur---more prominently \dfs---which are not parallelizable by \sys.

\heading{Results}

@@

\kk{I think the scoping of this should change to be both
  micro-benchmarks and comparison with existing work. Especially since
  we will have some form of comparison with Raftlib and phoenix. I
  guess this section is trying to establish \sys limits by comparing
  it with similar (but not identical) work.}

@@

\heading{Manual Parallelization}
Alternatively, a user might attempt to manually parallelize individual commands by using the abstractions provided by the POSIX shell.
The benefits depend on several factors, such as whether a command is I/O- or CPU-bound.

\heading{Streaming Framework}
In a different attempt to identify how far \sys is from ideal performance, we start from programs already written in a high-performance parallel framework and rewrite as shell programs.
Raftlib~\cite{} is a C++ library enabling streaming and dataflow parallelism without the fault-tolerance overheads of distributed systems.
We rewrite three Raftlib benchmarks---\ttt{wordcount}, \ttt{XX}, and \ttt{XX}---as shell scripts.
% These benchmarks were relatively easy found relatively easy to port and match many of the workloads one would expect to see in the shell as shell pipelines.

>>>>

\heading{Discussion}
\sys achieves near-linear speedup on low-parallelism configurations.
This is due to a high degree (deep trees) of task-based parallelism from \sys's multiple constructs---\eg \emph{fold}s and \ttt{relay}s---evident in all plots that contain \pur stages.
For example, Sort in $8\times$ spawns 37 nodes:
  8 \ttt{tr} nodes, 8 \ttt{sort} nodes, 7 \ttt{sort -m} nodes, and 14 \ttt{relay} nodes.
In effect, \sys hits the maximum number of parallel processes (and thus achieves optimal performance, \ie one without unnecessary process spawns and context switches) 
significantly earlier than the purported parallelism---$16-32\times$ for a 64-core system.
Asking for parallelism beyond that point degrades speedup, as overheads still increase but no further parallelism is available.
Sort-sort with $32\times$ has 314 nodes (sys: 54m) and with $64\times$ has 634 nodes (sys: 90m).
% 
%  On the other hand opt-bigrams 64 has 255 nodes and the sys time stays
%   the same so I think that it might have to do with the fact that
%   constant costs rise a lot (since opt-bigrams 32 takes just 1
%   minute).

@@

%% \TODO{An important point that needs to be made clear is that these are
%%   taken from the wild unmodified. And that Pash never leads to
%%   slowdown (except for the trivial head cases, where the slowdown is
%%   neligible).}
\heading{Results}
\sys's parallelism is set to $16\times$, as many of these pipelines are relatively long (\S\ref{ours}, Discussion).
Their input was generated by multiplying their original input several times, enough to reach 10GB.
Fig~\ref{fig:unix50-individual} shows the speedup (left) over the sequential runtime (right) for all 34 pipelines.
The average speedup is $5.49$ ($6.07\times$), and the weighted average (with the absolute times as weights) is $5.75\times$.
 % the \todo{geometric mean is $1.08\times$},

@@

\heading{Discussion}
Most pipelines see significant acceleration, except 13, 24, 25, 26, 29, 30 that see no speedup and 2, 19, 31 that see a slowdown.
The first group contain commands which \sys cannot parallelize---\eg \ttt{awk}, \ttt{sed} \ttt{-d}, and \ttt{tr} \ttt{-d}.
Some of them use general interpreters when they could be using a specialized equivalent tailored to the task at hand, the typical example being \ttt{awk} vs \ttt{cut}.
By applying only the replacements which to us experts felt more natural, we also give \sys the opportunity to identify and exploit parallelism---a common pattern we witnessed~\sx{macro1}.
For example, after this change, \todo{13} sees a \todo{XX}$\times$ speedup.

The second group contains \ttt{head}, thus practically processing only one line (regardless of the full input) and completing within $10ms$.
\sys's slowdown is due to constant setup costs of pipes and processes but execution still remains under $1s$.

@@

For the rest, \sys's speedup is \todo{capped} due to a combination of reasons:
  (i) they contain pure commands that are parallelizable but don't scale linearly, such as \ttt{sort} (0, 1, 3, 15, 16, 18, 20, 27, 28, 33),
  (ii) some are deep pipelines that already exploit task-based parallelism (7, 8, 9, 10, 11, 12, 14, 27, 28,) and
  (iii) they are not CPU-intensive, resulting in pronounced IO and constant costs (4, 5, 6, 7, 8, 12, 14, 22, 23).

@@

\heading{Take-aways} \sys accelerates unmodified pipelines found in the wild;
  small tweaks can yield further improvements, showing that \sys-awareness and scripting expertise can improve results.
Furthermore, \sys does not decelerate non-trivial computations, even when parallelism is impossible.

@@

\heading{Program}
This program is inspired by Hadoop's Definitive Guide~\cite[Chapter 2]{hadoop:15}, where it exemplifies a realistic mid-scale analytics pipeline and comprises three sub-tasks:
  download weather data from NCDC (shell), convert them to a Hadoop-friendly format (shell), and process them to get the maximum temperature (Hadoop).
Only the last of these tasks is the real focus of the book, % demonstrating the benefits of using distributed framework such as Hadoop.
  whereas we consider the end-to-end pipeline that implements all three sub-tasks---from fetching data to outputting the maximum temperature.
% The version used in the evaluation is slightly different than the one presented in Fig.~\ref{fig:example} (14 stages and totalling 240 characters), simplified only for clarity of exposition, in that it expresses the \ttt{for} loop as a first stage followed by \ttt{xargs}.

\heading{Results}
The original pipeline executes in \todo{XXs} on \todo{XX}GB of input data.
Applying \sys with \todo{XX}$\times$ parallelism leads to (XX speedup).
Naturally, different phases benefit differently from parallelization: \todo{part 1, 2, 3, 4}

It is worth reporting Hadoop's (3.2.1, Oracle JDK1.8.0\_255) execution time for the curious reader (see Discussion for fairness), which takes \todo{6m22s} (sequential) and \todo{4m30.554s} ($\times$\todo{XX}-parallel).
This amounts to \todo{XX}$\times$ speedup versus \todo{XX}$\times$ for \sys's equivalent (part 4).
% ---this includes downloading and uncompressing the data files which, as described earlier, is not expressible in the Hadoop equivalent.

In terms of program size, the Hadoop program amounts to 137 Java LoC as opposed to \todo{XX} shell LoC for \sys's \todo{part 4}.

@@

\heading{Discussion}
Similar to Unix50~\sx{unix50}, we found that large pipelines enable significant freedom in terms of expressiveness.
A few stages of the original script were expressed in a single \ttt{awk};
% unfortunately, \ttt{awk} is too general to have a single meaningful \todo{parallelizability} signature.
  a simpler (but longer) pipeline that leverages on \unix built-ins % might be outperformed by a specialized programs on a single node, but
  turns out to be trivially parallelizable.
% For example, the \ttt{sort} command at the end is an expensive version of extracting the maximum element;
%   this could have been expressed with a simple \ttt{awk} script that only keeps track of the max.

In terms of execution fairness, both systems are message-passing rather than shared-memory, and Hadoop's setup is atypical to improve fairness:
  it operates on non-HDFS file-system and has replication disabled.
Additionally, the Java program is written as a monolith tailored for this task---with ample opportunity for optimization from the compiler and runtime system---rather than a series of loosely coupled programs each written to handle many other cases; and the task at hand is an ideal fit for MapReduce system such as Hadoop.
Finally, execution time is long enough (\todo{XXh})to amortize Hadoop's startup costs.
% FIXME OSDI mention tabulation vs. sorting

\heading{Take-aways}
A multi-line parallel program specialized for the task at hand corresponds to a much smaller shell script with comparable performance (on a comparable setup).

@@

\heading{Program}
Coarsely, the pipeline is comprised of two distinct components % coarsely corresponding to a front- and a back-end
  communicating via \unix pipes.
The first downloads pages and extracts their URLs from their source, feeding back \todo{to} the start of the pipeline;
 the second extracts the text from HTML and applies natural-language processing---\eg trigrams, character conversion, term frequencies---to index it.
Collectively, they contains a total of 52 stages (first: 18; second: 34) written in multiple languages---\eg
its \ttt{url-extract}ion utility is written in JavaScript whereas its \ttt{word-stem}ming utility is in Python.
\sys can still operate on them as their parallelizability properties---\sta for \ttt{url-extract} and \ttt{word-stem}---can be trivially described by annotations.

\heading{Results}
% To simplify execution (and avoid Wikipedia), we have saved a recent Wikipedia version on a server within the network and to redirect all \ttt{curl} requests there.
\sys can operate on the entire program, but as its complexity makes end-to-end measurements difficult we report on the two components components individually.
With parallelism set to \todo{2/16} the execution time achieved by \sys is \todo{X} for the front-end and \todo{Y} for the back-end, while the original execution time is \todo{X} for the front-end and \todo{Y'} for the back-end.

%> \kk{If the front-end takes a lot of time (compared to the back-end), we might better not report it with the premise that it is a feedback cycle that cannot be parallelized by \sys.}

\heading{Discussion}
Since the front-end contains a feedback cycle, \sys \todo{doesn't
  parallelize it}, and in both cases it takes \todo{X seconds} to
complete. However, \sys parallelizes the back-end and achieves a
speedup of \todo{X} with parallelism set to \todo{2/16}. Several
stages of the pipeline are \sta (such as the \ttt{word-stem} script
and the text extraction) and that allows \sys to achieve benefits by
exposing data parallelism. Note that the back-end part of the script
contains 34 stages executing in parallel, so the sequential version
\todo{already benefits from task-based parallelism} \kk{not totally
  true}.

@@

% \subsection{Use Case: Container Orchestration}
% \label{macro3}
% 
% As a third realistic workload, we use a script for orchestrating containers to test the latest built across various Linux distributions.
% Compatibility testing is a common task for large software projects, commonly automated through shell scripting.
% 
% \heading{Program}
% The 30-line orchestration script is unusual among our set of benchmarks in that it is not structured as a linear pipeline.
% Rather, it makes extensive use of POSIX constructs common in the shell, including output redirection, conditionals, and here-documents.
% It also combines commands from other classes outside \sta and \pur---more prominently \dfs---which are not parallelizable by \sys.
% 
% \heading{Results}
% 
% \heading{Take-aways}

@@

%> \kk{I think the scoping of this should change to be both
%>   micro-benchmarks and comparison with existing work. Especially since
%>   we will have some form of comparison with Raftlib and phoenix. I
%>   guess this section is trying to establish \sys limits by comparing
%>   it with similar (but not identical) work.}

@@

%> \heading{Manual Parallelization}
%> Alternatively, a user might attempt to manually parallelize individual commands by using the abstractions provided by the POSIX shell.
%> The benefits depend on several factors, such as whether a command is I/O- or CPU-bound.
%> 
%> \heading{Streaming Framework}
%> In a different attempt to identify how far \sys is from ideal performance, we start from programs already written in a high-performance parallel framework and rewrite as shell programs.
%> Raftlib~\cite{} is a C++ library enabling streaming and dataflow parallelism without the fault-tolerance overheads of distributed systems.
%> We rewrite three Raftlib benchmarks---\ttt{wordcount}, \ttt{XX}, and \ttt{XX}---as shell scripts.
%> % These benchmarks were relatively easy found relatively easy to port and match many of the workloads one would expect to see in the shell as shell pipelines.
